Bayes network classifiers for fMRI

### 2012-11-09 article/article.tex
results

%% Abstract section.
\abstract{
Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain. This technique is typically used to determine which areas of the brain are involved in performing a certain task. Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data. In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data. By using classifiers we will try to predict the cognitive state of subjects, based on their brain activity during that time.
} % end of abstract

@@

\indent  The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time. The state that a person is in during some trial (e.g., looking at a face) is referred to as emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest. Specifically, the aforementioned studies try to infer the cognitive state of a subject from the brain activity that is evoked in a subject by performing that trial. For example, in an experiment where subjects look at a set of pictures (e.g., a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures. One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn for example what patterns of activity in the brain is associated with that cognitive state. This \emph{StarPlus} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures. These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}. \\
\indent In this study we use the same StarPlus data from the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences. An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes. However, for fMRI studies it is important that the pattern of activity that is identified by the algorithm is a pattern associated with the cognitive state. When using non-linear classifiers, it is possible to decode features predictive of the cognitive state, where the brain does not actually respond to these features. For example, when decoding the primary visual cortex with high enough spatial resolution, it would be possible to combine simple features with a non-linear neural network to decode the presence or absence of, for example, faces. However, this information is not actually processed in that part of the brain \cite{to:2012rev}. To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data. The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector (LSV) machines, and a Multivariate Gaussian (MV) classifier.
       Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\

@@

\section{Experimental Results}
\label{sec:results}
- raw results per algorithm

>>>>

%% Abstract section.
\abstract{Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain. This technique is typically used to determine which areas of the brain are involved in performing a certain task. Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data. In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data. By using classifiers we will try to predict the cognitive state of subjects, based on their brain activity during that time.
} % end of abstract

@@

\indent  The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time. The state that a person is in during some trial (e.g., looking at a face) is referred to as emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest. Specifically, the aforementioned studies try to infer the cognitive state of a subject from the brain activity that is evoked in a subject by performing that trial. For example, in an experiment where subjects look at a set of pictures (e.g., a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures. One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn for example what patterns of activity in the brain is associated with that cognitive state. This \emph{StarPlus} \cite{} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures. These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}. \\
\indent Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\

@@

\section{Experimental Results}
\label{sec:results}

Table~\ref{tab:results}

explain the symbols on the table, explain what rows (feature selection methods) are and what columns are (different classifiers and some of them with different structure on the data)

first column naive bayes with pca 148 components, others common feature selection methods as described in Section~\ref{sec:featureSelection}

* means different structure of the data

classifiers performance do not differ significantly

because even PCA has accuracy 61.5\% we can be sure that all the classifiers differ significantly from 50\%

mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)

\begin{table*}[htdp]
\centering
{\small
\begin{tabular}{c|c|c|c|c|c}
& \textbf{Naive Bayes} & \textbf{Naive Bayes} & \textbf{Linear SVC} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
PCA & 61.5 \% &  - & - & - & - \\
M-ROIs & - & 70.2 \% & 72.5 \% & 74.8 \% & 76.7 \% \\
Norm-M-ROIs & - & 71.7 \% & 71.3 \% & 75.8 \% & 77.1 \% \\
Norm-M-SD-ROIs & - & 65.6 \% & 69.8 \% & 69.4 \% & 65.0 \%\\
Norm/Vox-M-ROIs & - & 73.1 \% & 76.7 \% & 76.9 \% & 79.6 \% \\
\end{tabular}}
\caption{Performance results of three classifiers using different methods for feature selection and structure.}
\label{tab:results}
\end{table*}
